Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-13 aistats.tex
pass on optimization section

\paragraph{Optimization}
We can rewrite~\eqref{eq:defMAP} as
\alignn{
\m_n = \m_{n-1}- \lr_n (\m_{n-1} - T_n)
\label{eq:mean_update}
}
where $\lr_n := \inv{n_0 + n}$. 
Now let $f_X(\nat) = -\log p(X \cond \nat)$ such that $\E[f_X] = f$.
If we further introduce $g_n(\nat) := \nabla\logpart(\nat) - T_n = \nabla f_{X_n}(\nat)$, then~\eqref{eq:mean_update} becomes
\alignn{
\nabla\conj(\hat \nat_{n}) 
= \nabla\conj(\hat \nat_{n-1}) - \lr_n g_n(\hat \nat_{n-1})
}
which we recognize as the update formula for stochastic mirror descent (SMD) on $f$ with mirror map $\nabla\logpart$ and step-size schedule $\lr_n$.
\emph{MAP is an instantiation of SMD.}

@@

Surprisingly, we found no general solution to this seemingly simple problem, whether in the literature or by our own means.
There is a large body of work about asymptotic or large sample results, which we partially review in \S\ref{sec:insights}.
There are also results for special cases, such as $\cN(0, \sigma^2)$ (See \S\ref{sec:example}). 
But we are yet to find a solution encompassing a broad range of exponential families, and applicable to small sample sizes $n \lesssim d$.

{\bf Unboundedness of~\eqref{eq:bregmanMLE}.}
While~\eqref{eq:bregmanMAP} is always finite, \eqref{eq:bregmanMLE} may be infinite, 
for instance in Gaussian covariance estimation when $n \leq d + 1$.
Even worse, with categorical variables, there is always a non-zero probability to never sample one of the categories. In such the MLE will give zero weight to this category and $KL(p_{\nat^*} || p_\text{MLE} ) = +\infty$. 
This means that the expected KL~\eqref{eq:bregmanMLE} is infinite for any number of samples.
Instead of taking the expectation, one might want to bound the risk in high probability, without resorting to Markov inequality but this is a difficult endeavor.
These examples are rather making a case for regularized estimators such as MAP, for which we may find interesting upper bounds.

\section{Insights}
\label{sec:insights}

\subsection{Asymptote}
As a reference point for any finite convergence rate, it is interesting to know the asymptotic behavior of these quantities as $n \rightarrow +\infty$.
We now review classical asymptotic results. 
We point the reader towards \citet[\S1.1]{ostrovskii2021finite} for a more comprehensive overview.
Asymptotic results typically show that the MLE expressed with natural parameters $\nat$ is asymptotically normal: it reaches the Cramér-Rao lower bound \citep[for instance Ch4.2]{vdv1998asymptotic}.
With exponential families, the MAP is more simply expressed with mean parameters $\meanp$.
We get the asymptote by approximating the Bregman divergence using a second order Taylor expansion
\alignn{
    \bregmanconj(\m^* ; \m)
    &= \frac{\norm{\m^* - \m}^2_{\mF}}{2}
    + O(\norm{\m - \m^*}^3),
    \label{eq:bregman_asymptote}
}
where the Mahalanobis norm  $\| x \|_{\mF}^2 = x^\top \mF x$  is induced by $\mF  := \nabla^2\conj(\m^*)$, the hessian of the entropy at the optimum. It happens that  $\mF$ is also the inverse \textit{Fisher information matrix} at $\nat^*$, since
\aligns{
    \mF
    :=\nabla^2\conj(\m^*)
    = \nabla^2\logpart(\nat^*)^{-1}
    = \Cov_{\nat^*}[T(X)]^{-1}  \; .
}
Plugging~\eqref{eq:defMLE} or~\eqref{eq:defMAP} into~\eqref{eq:bregman_asymptote}, one can show that MLE and MAP verify 
\begin{align}
	\label{eq:MAP_asymptote}
	\E \bregmanconj \left (\E [T(X)] ; \hat \meanp_n^\text{MLE/MAP} \right )
	= \frac{d}{2n} + O(n^{- \frac{3}{2}}) \; .
\end{align}
See proof in the Appendix.
Both MLE and MAP have the same asymptote, as the contribution of the prior $n_0 \meanp_0$ gets negligible compared to the data for large $n$.
This asymptote is also independent of the optimum $\meanp^*$.

\subsection{Quadratic Case}
As another classical reference point, let us consider the case $\logpart(\nat) = \half \norm{\nat}_2^2$.
For instance, this is the log-partition of a Gaussian with variance $I$ and unknown mean 
\[ \cX=\real^d,\quad \nu(dx) = \exp(\half[-\|x\|^2])dx,\quad T(x)=x.\]
Then $\conj(\meanp) = \half \norm{\meanp}_2^2$ as well, and both Bregman divergences are squared $\ell^2$ distances:
\begin{align}
	\bregmanconj(\meanp^* ; \meanp) = \half \norm{\meanp^* -  \meanp }_2^2  \; .
\end{align}
Thanks to the independence of samples, we can break down the MLE into individual point's contributions
\begin{align}
	\expect{\half \norm{\m^* -  \inv{n}  \smallsum_i T_i}_2^2}
	=\frac{\Var(T)}{2n}
	=\frac{1}{2n}
\end{align}
Adding a reference mean $\m_0$ to get the MAP yields to
\begin{align}
	\expect{\bregmanconj(\meanp^*; \MAPm)}
	&= \frac{n \Var(T) +  n_0^2 \norm{\m^* -  \m_0}^2}{2(n+n_0)^2}.
	%\\ &= O\left(\frac{\Var(T)}{n} \right) + O\left(\frac{\norm{\m^* -  \m_0}^2}{n^2} \right)
	\label{eq:MAP_quadratic}
\end{align}
In both cases, we have a variance term in $O(n^{-1})$ and a bias term in $O(n^{-2})$. Ideally, such result should also similarly for arbitrary exponential families.
If we make restrictive assumptions on the log-partition function, we can relate $\bregmanconj$ to a norm or a quadratic.

{\bf If $\conj$ is $L$-Lipschitz} (e.g. $\logpart$ is defined within the $\ell^2$-ball of radius $L$), then
\begin{align}
    \bregmanconj(\m^* ; \m)
    &\leq L \norm{\m^* - \m} + \norm{\nat} \norm{\m^* - \m} \\
    &\leq 2L \norm{\m^* - \m}
\end{align}
so $\bregmanconj$ is $2L$-Lipschitz, and we get a $O(\inv{\sqrt{n}})$ rate.
Note that none of the standard exponential families verify this assumption.
\TODO{connection with optim non-smooth rate.}

@@

\section{Examples}
\label{sec:example}

@@

We now show how MAP can be interpreted as stochastic mirror descent (SMD).
This means that \textbf{1)} we may obtain a convergence rate for MAP from an optimization analysis, and \textbf{2)} any insights gained from the MAP may inform further design and analysis of SMD.
We then review the assumptions of relative smoothness, useful to deal with non-smooth functions, before investigating 3 analysis of SMD with the MAP.
\rlp{mention that examples have so far been lacking for relative SMD.}

\subsection{MAP as Stochastic Mirror Descent}
\label{ssec:MAP=SMD}
By Bayes rule, the posterior can always be written iteratively $p(\nat | \mD_n) \propto p(X_n | \nat) p(\nat | \mD_{n-1})$.
Plugging in~\eqref{eq:joint_likelihood}, and introducing $f_n(\nat) = - \log p(X_{n} | \nat)$, we see that the MAP verifies
\begin{align*}
	\hat \nat_{n}
	= \argmin_\nat f_n(\nat) + (n_0 + n -1) \bregman(\nat; \hat \nat_{n-1})
    %\label{eq:SBPP}
\end{align*}
which is the update of a Stochastic Bregman Proximal Point update with step-size $\inv{n_0 + n - 1}$.
%Maybe skip this SBPP story altogether
Expanding $f_n(\nat) = \logpart(\nat) - T_n$ and  playing with the Bregman we get
\begin{align*}
		\hat \nat_{n}
	= \argmin_\nat \langle g_{n}(\hat \nat_{n-1}), \nat \rangle + (n_0 + n) \bregman(\nat; \hat \nat_{n-1})
    %\label{eq:SMD}
\end{align*}
where $g_{n}(\nat) := \nabla f_n(\nat) = \hat \m_{n-1} - T_n$.
We recognize the update of Stochastic Mirror Descent (a.k.a. Stochastic Bregman Gradient) with step-size $\lr_n = \inv{n_0 + n}$.
From the first order optimality condition, we see this is a stochastic gradient step in the dual $\hat \m_n = \hat \m_{n-1} - \lr_n g_n$.
This parallel means we could obtain a finite convergence rate for the  MAP from an analysis of these algorithms.
In particular, it could potentially apply to a wide range of distribution $\cD$ on $X$.
Note that this optimization perspective does not hold for the MLE %($n_0=0$)
as its implicit uniform prior may correspond to an initialization outside of $\Theta$.

\subsection{Relative Smoothness}
Mirror descent (MD)\footnote{
Also known as
Bregman (proximal) gradient, relative gradient descent or NoLips.
}
\citep{nemirovski1983problem,beck2003mirror}
%Nemirovski introduced the algorithm, and Beck established the connection with Bregman divergences, e.g. the argmin form.
and SMD 
\citep{nemirovski2009robust,ghadimi2012optimal} 
are typically encountered in non-smooth (online) optimization,
under bounded (or Lipschitz) gradient assumption on the objective $f$
and strong convexity assumption on the potential $\logpart$
\citep[Th.4.2 \& Th.6.3]{bubeck2015convex}.
In our case, none of these assumptions may hold. 
For instance $\logpart = -\log$ is neither smooth nor strongly convex.

Recently, these assumptions have been relaxed to allow for non-smooth functions $f$
\footnote{$f$ is $L$-smooth if its gradient is $L$-Lipschitz}.
and non-strongly convex potential $\logpart$, first in the deterministic setting
\citep{birnbaum2011distributed, bauschke2017descent, lu2018relatively}, then in the stochastic setting \citep{hanzely2018fastest, dragomir2021fast, dorazio2021stochastic}.
These assumptions have been replaced with
the $\mu$-strong convexity and $L$-smoothness of $f$
\emph{relative} to $\logpart$, defined as
\aligns{
	\mu \cB_{A}(x, y)
	\leq
	\cB_f(x,y)
	\leq
	L \cB_A(x,y) \; .
}
When $\logpart = \norm{\cdot}^2$, we recover standard (Euclidean) smoothness, and gradient descent. 
Those conditions ensure the linear convergence rate of MD with $A$ as the reference function.
Exponential families log-likelihood perfectly fit into this framework, as
\aligns{
	f(\theta) = A(\theta) - \expect{\lin{X, \theta}}
}
is $1$-smooth and $1$-strongly convex relative to $A$.

\subsection{Bounding the Randomness}
Smoothness is not the only issue.
To analyze stochastic algorithms, one also need to quantify the randomness.
Many assumptions exist for SGD \citep[\S3 for a modern review]{khaled2020better}.
We now review attempts at analyzing SMD under relative smoothness assumption, in the light of  the MAP example.
See \cref{tbl:assumptions} for a summary.

@@

\subsubsection{Bounded Variance}
\citet{hanzely2018fastest} assume the covariance between negative gradient and next iterate is upper bounded at every time step: $\forall n \geq 1, \forall \hat \nat_{n-1},$
\alignn{
\Cov(-g_n(\hat \nat_{n-1}) , \MAPt) \leq \lr_n C
\label{eq:hanzely}
}
for some constant $C$.
For a Gaussian with known variance ($A(\theta) = \frac{1}{2}\norm{\theta}^2$),
this definition recovers the variance of the stochastic gradient
$\expect[\tilde g_t\!\!]{\norm{\nabla f(\theta) - \tilde g(\theta)}{}^2}\leq C$.
More generally, if $A$ is $\mu$-strongly convex in $\norm{\cdot}$,
the LHS of \cref{eq:hanzely} is bounded by
$\expecti{\frac{1}{\mu}\norm{\nabla f(\theta) - \tilde g(\theta)}{}^2_*}$,
and a bound on the variance of the stochastic gradient is a sufficient condition for \cref{eq:hanzely}
to hold.
Another perspective on this covariance is as the expectation of the symmetrized Bregman $\cS_\conj(\m_1, \m_2) = \bregmanconj(\m_1, \m_2) + \bregmanconj(\m_2, \m_1)$ between stochastic and deterministic updates
\aligns{
\Cov(-g_n(\hat \nat_{n-1}) , \MAPt) = \E[\cS(\MAPm - \lr g_n  )]
}
thus showing positivity.
Under this assumption, they prove a $O(1/n)$ convergence rate with a $O(1/n)$ step-size \citep[Lem.4.8]{hanzely2018fastest}, by applying a polynomial tail averaging \citep{lacostejulien2012simpler} in primal space $\Theta$.

@@

\citet{dragomir2021fast} introduce a much weaker assumption $\forall n \geq 1, \forall \MAPm,$
\alignn{
	\expect[\tilde g]{
		\cB_{A^*}(\MAPm - 2\lr g(\theta_*), \MAPm)
	} \leq 2 \lr^2 C \; .
	\label{eq:dragomir}
}
In words, this is bounding the variance of the gradients at the minimum, but with a metric that depends on the iterate $\MAPm$, and this bound should hold for every iterate we may encounter on the stochastic trajectory.
% As observed by Frederick, this is really weaker than Hanzely, because it is only the first half of their assumption. There is only a factor 2 that's preventing a direct inequality. 
Under this assumption, they prove convergence up to a variance ball with a constant step-size.
Their descent lemma \citep[Eq.(12)]{dragomir2021fast} is the exact analog  of a modern analysis of SGD \citep[Th.3.2]{gower2019sgd}, meaning we can plug in a $O(1/n)$ decreasing step-size and obtain a $O(1/n)$ convergence rate.

Seems like a perfect fit then ? Well not really, this variance assumption does not hold for exactly the same reason as the one of Hanzely. Having something valid for every possible iterates is way too strong.

\subsubsection{Bounded Optimality Gap}
Inspired by \citet{loizou2021stochastic}, \citet{dorazio2021stochastic} explores the hypothesis
\alignn{
\min_\nat f(\nat) - \expect[X]{\min_\nat f_X(\nat)} \leq C 
\label{eq:dorazio}
}
where $f_X$ is a stochastic estimate of $f = \expecti{f_X}$. In our case $f_X(\nat) = - \log p(X\cond \nat)$.
In words, this is lower bounding the expectation of the minimum of these stochastic estimates.
For probabilistic models, such a bound is finite as soon as the model cannot give infinite density to any data point $x$.
This holds for instance for discrete distributions 
-- $f_x\geq 0$ because the probability mass is upper bounded by $1$ -- 
but it rules out  many families.
For normal distributions set $\mu=x$ and $\sigma^2 \rightarrow 0$ to get $p(x\cond \nat) \rightarrow +\infty$. 
For gamma, set $\alpha = \beta x$ and $\beta \rightarrow +\infty$.
For beta, set $\alpha=\beta \frac{x}{1-x}$ and $\beta \rightarrow +\infty$.
 Other examples include inverse Gaussians, log-normal, gamma, inverse gamma.

 It is possible to overcome this limitation though, by treating batches of samples as 1 sample, and averaging sufficient statistics -- e.g. $Y = \{X_1, \dots, X_k\}$ and $T(Y) = \inv{k}\sum_i T(X_i)$.
 For instance a multivariate normals of dimension $d$ cannot attribute infinite density to $d+1$ samples that are not in an affine subspace.

So it seems this approach can yield interesting convergence rates for the MAP, and they do prove linear convergence of the last iterate up a to a variance ball with constant step-size, but the size of this ball $\frac{C}{\alpha}$ \rlp{strong convexity notation} is independent of  the step-size. Their descent lemma does not lead to convergence with a decreasing $O(1/t)$ step-size.

>>>>

\paragraph{Optimization}
We can rewrite~\eqref{eq:defMAP} as
\alignn{
\m_n = \m_{n-1}- \lr_n (\m_{n-1} - T_n)
\label{eq:mean_update}
}
where $\lr_n := \inv{n_0 + n}$. 
Now let $f_X(\nat) = -\log p(X \cond \nat)$ such that $\E[f_X] = f$.
If we further introduce $g_n(\nat) := \nabla\logpart(\nat) - T_n = \nabla f_{X_n}(\nat)$, then~\eqref{eq:mean_update} becomes
\alignn{
\nabla\conj(\hat \nat_{n}) 
= \nabla\conj(\hat \nat_{n-1}) - \lr_n g_n(\hat \nat_{n-1})
}
which we recognize as the update formula for stochastic mirror descent (SMD) on $f$ with mirror map $\nabla\logpart$ and step-size schedule $\lr_n$.

@@

Surprisingly, we found no general solution to this seemingly simple problem, whether in the literature or by our own means.
%TODO. what we did and are going to present next.
We are yet to find a solution encompassing a broad range of exponential families, and applicable to small sample sizes $n$.

{\bf Remark.}
While~\eqref{eq:bregmanMAP} is always finite, \eqref{eq:bregmanMLE} may be infinite, 
for instance in Gaussian covariance estimation when $n \leq d + 1$.
Even worse with categorical variables, the MLE is always ill-defined. %TODO
Instead of taking the expectation, one might want to bound the risk in high probability, without resorting to Markov inequality.

\section{Insights}

\subsection{Asymptote}
As a reference point for any finite convergence rate, it is interesting to know the asymptotic behavior of these quantities as $n \rightarrow +\infty$.
Asymptotic results typically show that the MLE expressed with natural parameters $\nat$ is asymptotically normal: it reaches the Cramér-Rao lower bound \citep[for instance Ch4.2]{vdv1998asymptotic}.
With exponential families, the MAP is more simply expressed with mean parameters $\meanp$.
We get the asymptote by approximating the Bregman divergence using a second order Taylor expansion
\begin{align}
    \bregmanconj(\m^* ; \m)
    &= \frac{\norm{\m^* - \m}^2_{\mF}}{2}
    + O(\norm{\m - \m^*}^3),
\end{align}
where the Mahalanobis norm  $\| x \|_{\mF}^2 = x^\top \mF x$  is induced by $\mF  := \nabla^2\conj(\m^*)$, the hessian of the entropy at the optimum. It happens that  $\mF$ is also the inverse \textit{Fisher information matrix}, since
\begin{align}
    \mF
    :=\nabla^2\conj(\m^*)
    = \nabla^2\logpart(\nat^*)^{-1}
    = \Cov_{\nat^*}[T(X)]^{-1}  \; .
\end{align}
From there, it is possible to show an asymptotic rate for the MLE and the MAP \textcolor{red}{D.S.: How?}
\begin{align}
\label{eq:MAP_asymptote}
	\E \bregmanconj \left (\E [T(X)] ; \hat \meanp_n^\text{MLE/MAP} \right )
	= \frac{d}{2n} + O(n^{- \frac{3}{2}}) \; .
\end{align}
Both MLE and MAP have the same asymptote, as the contribution of the prior $n_0 \meanp_0$ gets negligible compared to the data for large $n$.
This asymptote is also independent of the optimum $\meanp^*$.
We point the curious reader towards \citet[\S1.1]{ostrovskii2021finite} for a more comprehensive review.

\subsection{Quadratic Case}
\rlp{make it multivariate, with a reference to \citep{bach2013nonstronglyconvex}?}
As another reference point, let us consider the case $\logpart(\nat) = \half \norm{\nat}_2^2$.
For instance, this is the log-partition of a Gaussian with variance $1$ and unknown mean 
\[ \cX=\real,\quad \nu(dx) = \exp(\half[-x^2])dx,\quad T(X)=X.\]
Then $\conj(\meanp) = \half \norm{\meanp}_2^2$ as well, and both Bregman divergences are squared $\ell^2$ distances:
\begin{align}
	\bregmanconj(\meanp^* ; \meanp) = \half \norm{\meanp^* -  \meanp }_2^2  \; .
\end{align}
Thanks to the independence of samples, we can break down the MLE into individual point's contributions
\begin{align}
	\expect{\half \norm{\m^* -  \inv{n}  \smallsum_i T_i}_2^2}
	=\frac{\Var(T)}{2n}
	=\frac{1}{2n}
\end{align}
Adding a reference mean $\m_0$ to get the MAP yields to
\begin{align}
	\expect{\bregmanconj(\meanp^*; \MAPm)}
	&= \frac{n \Var(T) +  n_0^2 \norm{\m^* -  \m_0}^2}{2(n+n_0)^2}.
	%\\ &= O\left(\frac{\Var(T)}{n} \right) + O\left(\frac{\norm{\m^* -  \m_0}^2}{n^2} \right)
	\label{eq:MAP_quadratic}
\end{align}
In both cases, we have a variance term in $O(n^{-1})$ and a bias term in $O(n^{-2})$. Ideally, such result should also similarly for arbitrary exponential families.
If we make restrictive assumptions on the log-partition function, we can relate $\bregmanconj$ to a norm or a quadratic.

{\bf If $\conj$ is $L$-Lipschitz} (e.g. $\logpart$ is defined within the $\ell^2$-ball of radius $L$), then
\begin{align}
    \bregmanconj(\m^* ; \m)
    &\leq L \norm{\m^* - \m} + \norm{\nat} \norm{\m^* - \m} \\
    &\leq 2L \norm{\m^* - \m}
\end{align}
so $\bregmanconj$ is $2L$-Lipschitz, and we get a $O(\inv{\sqrt{n}})$ rate.
\TODO{connection with optim non-smooth rate.}

@@

\section{Examples}

@@

This section show how MAP can be interpreted as stochastic mirror descent (SMD). Thanks to this strong link, we show that \textbf{1)} we may obtain a convergence rate for MAP from the rate od SMD, and \textbf{2)} any insights gained from the MAP may inform on further design and analysis of SMD.
Then, we review the assumptions of relative smoothness, useful to deal with non-smooth functions, before investigating three different analyses of SMD with the MAP.
\rlp{mention that examples have so far been lacking for relative SMD.}

\subsection{MAP as Stochastic Mirror Descent}
\label{ssec:MAP=SMD}
By the Bayes rule, the posterior can always be written iteratively $p(\nat | \mD_n) \propto p(X_n | \nat) p(\nat | \mD_{n-1})$.
Plugging in~\eqref{eq:joint_likelihood}, and introducing $f_n(\nat) = - \log p(X_{n} | \nat)$, we see that the MAP verifies
\begin{align*}
	\hat \nat_{n}
	= \argmin_\nat f_n(\nat) + (n_0 + n -1) \bregman(\nat; \hat \nat_{n-1})
    %\label{eq:SBPP}
\end{align*}
which is the update of a Stochastic Bregman Proximal Point update with step-size $\inv{n_0 + n - 1}$.
%Maybe skip this SBPP story altogether
If we expand $f_n(\nat) = \logpart(\nat) - T_n$,  we get (after some manipulation with the Bregman divergence)
\begin{align*}
		\hat \nat_{n}
	= \argmin_\nat \langle g_{n}(\hat \nat_{n-1}), \nat \rangle + (n_0 + n) \bregman(\nat; \hat \nat_{n-1})
    %\label{eq:SMD}
\end{align*}
where $g_{n}(\nat) := \nabla f_n(\nat) = \hat \m_{n-1} - T_n$.
We recognize the update of Stochastic Mirror Descent (a.k.a. Stochastic Bregman Gradient) with step-size $\lr_n = \inv{n_0 + n}$.
From the first order optimality condition, we see this is a stochastic gradient step in the dual $\hat \m_n = \hat \m_{n-1} - \lr_n g_n$.
This parallel means we could obtain a finite convergence rate for the  MAP from an analysis of these algorithms.
In particular, it could potentially apply to a wide range of distribution $\cD$ on $X$.
Note that this optimization perspective does not hold for the MLE %($n_0=0$)
as its implicit uniform prior may correspond to an initialization outside of $\Theta$.

\subsection{Relative Smoothness}
Mirror descent (MD) \citep{nemirovski1983problem,beck2003mirror}, also known as
Bregman (proximal) gradient, relative gradient descent or NoLips,
%Nemirovski introduced the algorithm, and Beck established the connection with Bregman divergences, e.g. the argmin form.
and SMD 
\citep{nemirovski2009robust,ghadimi2012optimal} 
are typically encountered in non-smooth (online) optimization,
under bounded (or Lipschitz) gradient assumption on the objective $f$
and strong convexity assumption on the potential $\logpart$
\citep[Th.4.2 \& Th.6.3]{bubeck2015convex}.
In our case, none of these assumptions may hold. 
For instance $\logpart = -\log$ is neither smooth nor strongly convex.

Recently, these assumptions have been relaxed to the $\mu$-strong convexity and $L$-smoothness of $f$
\emph{relative} to $\logpart$, defined as
\aligns{
	\mu \cB_{A}(x, y)
	\leq
	\cB_f(x,y)
	\leq
	L \cB_A(x,y) \; .
}
This extends the theory to a family of non-smooth functions $f$ and non-strongly convex potential $\logpart$, first in the deterministic setting
\citep{birnbaum2011distributed, bauschke2017descent, lu2018relatively}, then in the stochastic setting \citep{hanzely2018fastest, dragomir2021fast, dorazio2021stochastic}. When $\logpart = \norm{\cdot}^2$, we recover standard (Euclidean) smoothness, and gradient descent. 
Those conditions ensure the linear convergence rate of MD with $A$ as the reference function.

\textcolor{red}{Sure of this formulation?} Exponential families log-likelihood perfectly fit into this framework, as
\aligns{
	f(\theta) = A(\theta) - \expect{\lin{X, \theta}}
}
is $1$-smooth and $1$-strongly convex relative to $A$. However, the (relative) smoothness is not biggest issue in our convergence analysis, but the bound on the variance, as explained in the next section.

\subsection{Bounding the Randomness}
To analyze stochastic algorithms, one also need to quantify the randomness.
Many assumptions exist for SGD \citep[\S3 for a modern review]{khaled2020better}.
We now review attempts at analyzing SMD under relative smoothness assumption, in the light of  the MAP example.
See \cref{tbl:assumptions} for a summary.

@@

\subsubsection{Bounded Variance}
\citet{hanzely2018fastest} assume the covariance between negative gradient and next iterate is upper bounded at every time step: $\forall n \geq 1, \forall \hat \nat_{n-1},$
\alignn{
\Cov(-g_n(\hat \nat_{n-1}) , \MAPt) \leq \lr_n C
\label{eq:hanzely}
}
for some constant $C$.
For a Gaussian with known variance ($A(\theta) = \frac{1}{2}\norm{\theta}^2$),
this definition recovers the variance of the stochastic gradient
\[
	\expect[\tilde g_t\!\!]{\norm{\nabla f(\theta) - \tilde g(\theta)}{}^2}\leq C.
\]
More generally, if $A$ is $\mu$-strongly convex w.r.t. $\norm{\cdot}$,
the LHS of \cref{eq:hanzely} is bounded by
$\expecti{\frac{1}{\mu}\norm{\nabla f(\theta) - \tilde g(\theta)}{}^2_*}$,
and a bound on the variance of the stochastic gradient is a sufficient condition for \cref{eq:hanzely}
to hold.
\textcolor{red}{check formulation} Another perspective on this covariance is as the expectation of the symmetrized Bregman $\cS_\conj(\m_1, \m_2) = \bregmanconj(\m_1, \m_2) + \bregmanconj(\m_2, \m_1)$ between stochastic and deterministic updates
\aligns{
\Cov(-g_n(\hat \nat_{n-1}) , \MAPt) = \E[\cS(\MAPm - \lr g_n  )]
}
thus showing positivity.
Under this assumption, \textcolor{red}{who is they?}they prove a $O(1/n)$ convergence rate with a $O(1/n)$ step-size \citep[Lem.4.8]{hanzely2018fastest}, by applying a polynomial tail averaging \citep{lacostejulien2012simpler} in primal space $\Theta$.

@@

\citet{dragomir2021fast} introduce a much weaker assumption $\forall n \geq 1, \forall \MAPm,$
\alignn{
	\expect[\tilde g]{
		\cB_{A^*}(\MAPm - 2\lr g(\theta_*), \MAPm)
	} \leq 2 \lr^2 C \; .
	\label{eq:dragomir}
}
In words, this bounds the variance of the gradients at the minimum, but with a metric that depends on the iterate $\MAPm$, and this bound should hold for every iterate we may encounter on the stochastic trajectory.
% As observed by Frederick, this is really weaker than Hanzely, because it is only the first half of their assumption. There is only a factor 2 that's preventing a direct inequality. 
Under this assumption, they prove convergence up to a variance ball with a constant step-size.
Their descent lemma \citep[Eq.(12)]{dragomir2021fast} is the exact analog  of a modern analysis of SGD \citep[Th.3.2]{gower2019sgd}, meaning we can plug in a $O(1/n)$ decreasing step-size and obtain a $O(1/n)$ convergence rate.

\textcolor{red}{Too informal, plz rephrase} Seems like a perfect fit then ? Well not really, this variance assumption does not hold for exactly the same reason as the one of \textcolor{red}{use cite command} Hanzely. Having something valid for every possible iterates is way too strong.

\subsubsection{Bounded Optimality Gap}
Inspired by \citet{loizou2021stochastic}, \citet{dorazio2021stochastic} explore the hypothesis
\alignn{
\min_\nat f(\nat) - \expect[X]{\min_\nat f_X(\nat)} \leq C,
\label{eq:dorazio}
}
where $f_X$ is a stochastic estimate of $f = \expecti{f_X}$. In our case $f_X(\nat) = - \log p(X\cond \nat)$.
In words, this lower bounds the expectation of the minimum of the stochastic estimates.
For probabilistic models, such a bound is finite as soon as the model cannot give infinite density to any data point $x$.
This holds for instance for discrete distributions because the probability mass is upper bounded by $1$,
but it rules out  many families.
In the case of normal distributions, setting $\mu=x$ and $\sigma^2 \rightarrow 0$ gets $p(x\cond \nat) \rightarrow +\infty$. \textcolor{red}{Check this, I changed the formulation but i am not sure if this remains correct or not.} We have a similar behavior for gamma distribution with $\alpha = \beta x$ and $\beta \rightarrow +\infty$, or with the beta distribution with $\alpha=\beta \frac{x}{1-x}$ and $\beta \rightarrow +\infty$.
 Other examples include inverse Gaussians, log-normal, gamma, inverse gamma.

It is possible to overcome this limitation by treating batches of samples as single samples by averaging sufficient statistics, e.g.,. $Y = \{X_1, \dots, X_k\}$ and $T(Y) = \inv{k}\sum_i T(X_i)$.
For instance, a multivariate normals of dimension $d$ cannot attribute infinite density to $d+1$ samples that are not in an affine subspace.

In conclusion, it seems these approaches can yield to interesting convergence rates for the MAP, and they do prove linear convergence of the last iterate up a to a variance ball with constant step-size, but the size of this ball $\frac{C}{\alpha}$ \rlp{strong convexity notation} is independent of the step-size. \textcolor{red}{Who is their?} Their descent lemma does not lead to convergence with a decreasing $O(1/t)$ step-size.
